### 2020-05-22 production/prod_analysis-cfa.rmd

---
title: "Meritocracy and distributive preferences ISSP"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
css: "../input/css/custom.css"
output:
  html_document:
    number_sections: true
    toc: true
    toc_float:
        collapsed: false
    toc_depth: 2
    code_folding: "hide"
editor_options:
  chunk_output_type: console
---
Hacer una descripcion de la linea de trabajo en general.

```{r eval=FALSE, include=FALSE}
rmarkdown::render(input = "production/prod_analysis-cfa.rmd",output_format = "html_document"); browseURL(url = "production/prod_analysis-cfa.html")
```

```{r include=FALSE}
knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = TRUE)
knitr::opts_knit$set(root.dir = "../")
options(scipen=999)
rm(list=ls())
options(knitr.kable.NA = '')
```

**Libraries**
```{r}
library(sjPlot)
library(sjmisc) # several functions,ej read_stata
library(descr)
library(memisc)  # codebook
library(psych) # cor2latex
library(corrplot) # graph correlations
library(lavaan)
library(stargazer) 
library(semTools)
library(plyr) # count
library(dplyr)
#library(car) # recode
library(ggplot2)
#library(Hmisc)
#library(tidyverse)
library(data.table)
library(stringr)
library(countrycode)
library(lme4)
library(texreg)
library(mice)
library(ggrepel) # for scatter labels in sjp.scatter	
#library(semPlot)
library(haven)
library(sjlabelled) 
library(kableExtra)

```

**Load data**

```{r}
load(file = "input/data/proc/redistrib.rda")
```

# Descriptives
```{r results='asis'}
  # Summary table
    stargazer(redistrib[-c(1,2)], type="html")

  # Summary table to latex
    #sink("output/results/desctab_red.txt")
    #stargazer(redistrib[-c(1,2)], type="text"); sink()

  # Simple graph:
    sjPlot::plot_frq(redistrib$income_dif, show.values = TRUE, coord.flip = TRUE)

```

```{r echo=FALSE}
# Likert graph 
    sjPlot::plot_frq((redistrib[3:7])) # ajustar

    # short var labels
    items <- c("Gov. reduce income dif.", "Gov. decent unemp.",
               "High income taxes", "Buy better health", 
               "Buy better education")

    # cambiar orden para la leyenda  
    labels= c("Strongly agree", "Agree", "Disagree",
              "Strongly disagree", "(neither/nor)" )  ### E

    # ajustes globales sjplot
    sjPlot::set_theme(theme         = "blank",
                      legend.inside = TRUE, 
                      legend.pos    = "bottom right", legend.size = 0.8,
                      geom.outline.size=0, geom.label.size =3)
    #   sjp.setTheme(theme         = "blank",
    #             legend.inside = TRUE, 
    #            legend.pos    = "bottom right", legend.size = 0.8,
    #           geom.outline.size=0, geom.label.size =3)

    sjPlot::plot_likert((redistrib[3:7]), 
                        axis.labels   = items,
                        legend.labels = labels,
                        cat.neutral   = 5, # identifica a indiferentes
                        geom.colors   = c("#9ecae1", "#6baed6","#4292c6", 
                                          "#2171b5"), # colorbrewer2.org 
                        sort.frq      = "neg.asc", # sort descending)
                        title         = "Agree / disagree ...",
                        intercept.line.color = "white", # vertical middle
                        expand.grid   = FALSE, # no inner margins in plot
                        show.n        = FALSE, # hide N's in axis labels
                        grid.range    = 1.40) # expand axis 130% )

    # sjp.likert((redistrib2[3:7]), 
    #             axis.labels   = items,
    #            legend.labels = labels,
    #           cat.neutral   = 3, # identifica a indiferentes
    #          geom.colors   = c("#9ecae1", "#6baed6","#4292c6", 
    #                           "#2171b5"), # colorbrewer2.org 
    #        sort.frq      = "neg.asc", # sort descending)
    #       title         = "Agree / disagree ...",
    #     expand.grid   = FALSE, # no inner margins in plot
    #     show.n        = FALSE, # hide N's in axis labels
    #    grid.range    = 1.40) # expand axis 130%

    # Save
    dev.copy(png,"output/results/likert_red.png",width=800); dev.off()

```

```{r}
# 2.3 Measurement structure ------------- ########

# Estimación del modelo

	load("input/data/proc/redistrib.rda")

	cfa_red1 <- '
	# latent variables
	redist =~ income_dif + unemployed + taxes 
	accum  =~ health + educ'

	fit_red1 <- cfa(cfa_red1,data=redistrib, missing = "ML") 
#	summary(fit_red1, fit.measures=TRUE,standardized=TRUE)

###	################## 	agregar un sumary mas  reducido

		# Path

	#		semPlot::semPaths(fit_red1,layout = "tree2", rotation = 2, 
	#  	intercepts = FALSE,	style = "lisrel", curvePivot = TRUE,
	#  	cardinal = TRUE)
	#	  dev.copy(png,"output/results/path_red1.png",width=750,height=600)
	#	  dev.off()	  	  	 
## ## ## ## ### ##                                            ### ## ## ## ##ARREGLAR LOS DIAGRAMAS. 
		# Path - with results
	#	  semPlot::semPaths(fit_red1,"std",layout = "tree2", rotation = 2, 
	#  	intercepts = FALSE,	style = "lisrel", curvePivot = TRUE,
	#  	cardinal = TRUE, edge.color = "black", residScale = 10)
	#	  dev.copy(png,"output/results/path_red1_res.png",
	#	  	width=750,height=600); dev.off()

	# Categorical	  
		fit_red1c <- cfa(cfa_red1,data=redistrib, 
			ordered = c("income_dif","unemployed","taxes", "health", 
			"educ")) 
	#	summary(fit_red1c, fit.measures=TRUE,standardized=TRUE)
###	################## 	                                         agregar un sumary mas  reducido

sum_fit<- bind_rows(fitmeasures(fit_red1)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")])

sum_fit$mod <- c("Model 1")
sum_fit$est <- c("MLR")
#sum_fit <- select(sum_fit,mod,est,everything())
colnames <- c("Model","Estimator","$\\chi^2$","df","CFI","CFI.sca","CFI.rob","RMSEA","RMSEA.sca","RMSEA.rob")

sumtable01<- kable(sum_fit,digits = 3,format = "html",row.names = F,booktabs=T, caption = "Summary fit indices",col.names = colnames,escape = FALSE) %>%
  kable_styling(full_width = F)  %>%
      collapse_rows(columns = 1,valign = "middle")  %>%
      footnote(number = c("Model 1: fixed order by percepction/preferenc"))
sumtable01

		# 2.4 Factor scores --------------------- --------

	# Multiple imputation (para recuperar missings en scores)

	  # generate 5 multiple complete datasets 
		out <- mice(redistrib[3:7], m=5) 
		D1 <- complete(out, 1) 
		D2 <- complete(out, 2) 
		D3 <- complete(out, 3) 
		D4 <- complete(out, 4) 
		D5 <- complete(out, 5)

		# fit model for each complete dataset 
		fit1 <- cfa(cfa_red1, data=D1) 
		fit2 <- cfa(cfa_red1, data=D2) 
		fit3 <- cfa(cfa_red1, data=D3) 
		fit4 <- cfa(cfa_red1, data=D4) 
		fit5 <- cfa(cfa_red1, data=D5)

	# predict scores for all models 
		p1 <- predict(fit1) 
		p2 <- predict(fit2) 
		p3 <- predict(fit3) 
		p4 <- predict(fit4) 
	p5 <- predict(fit5)

# compute 'average' across 5 sets of scores: 
		scores <- (p1 + p2 + p3 + p4 + p5)/5

 #Generate factor scores      
   scores_dat=as.data.frame(scores)
    dim(scores_dat)
    str(scores_dat)

# Adaptar base para agregar factor scores (sacar todos NA)  
    dim(redistrib)
   redistrib_m=redistrib[rowSums(is.na(redistrib[3:7]))!=5,]
    dim(redistrib_m)

# Merge with factor scores
	  redistrib_sco=cbind(redistrib,scores)

	  # Check
	    summary(redistrib_sco$redist)
	    summary(redistrib_sco$accum)

	# Save  factor scores
#	  save(redistrib_sco, file="input/data/proc/redistrib_sco.rda")

```

![IMAGEN PROVISORIA](../output/results/path_red1_res.png)
```{r}
# 4. MODELS A:  MERIT & REDISTRIB ========= =

# 4.1 Merge objects --------------------- ---- 
	load("input/data/proc/redistrib.rda")
	load("input/data/proc/getahead.rda")                       
	load("input/data/proc/detearn.rda")

	aux1 <- merge(getahead,redistrib,by=c("id","v5"))
  merit_redist=merge(aux1,detearn,by=c("id","v5"))

# 4.2 Measurement & structural model ---- --------

  # Measurement simple 
	  cfa_mer_red1 <- '
	  # latent variables
		  merit =~ hwork + ambition 
		  parent=~ wealthy + pareduc
		  backg=~ race + gender
			network=~ people + polcone
		  bmerit =~ welljob + hardjob 
		  need=~ family + child
			redist =~ income_dif + unemployed + taxes 
	  	accum  =~ health + educ'

	  fit_mer_red1 <- cfa(cfa_mer_red1,data=merit_redist, 
	  	missing = "ML")

#	 summary(fit_mer_red1, fit.measures=TRUE,standardized=TRUE)

	  	# Path
	 		#  semPlot::semPaths(fit_mer_red1,layout = "circle", rotation = 3, 
			#  	intercepts = FALSE, style = "lisrel", curvePivot = TRUE,
			#  	sizeMan = 4, sizeLat = 6)
	 		#  dev.copy(png,"output/results/path_mer_red1.png",width=800,
		  #  	height=650);  dev.off()

	  fit_mer_redc <- cfa(cfa_mer_red1,data=merit_redist,
	  	ordered = names(merit_redist[3:10,19:29]))
	#	summary(fit_mer_redc, fit.measures=TRUE,standardized=TRUE)

sum_fit<- bind_rows(fitmeasures(fit_mer_redc)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")])

sum_fit$mod <- c("Model 1")
sum_fit$est <- c("MLR")
#sum_fit <- select(sum_fit,mod,est,everything())
colnames <- c("Model","Estimator","$\\chi^2$","df","CFI","CFI.sca","CFI.rob","RMSEA","RMSEA.sca","RMSEA.rob")

sumtable01<- kable(sum_fit,digits = 3,format = "html",row.names = F,booktabs=T, caption = "Summary fit indices",col.names = colnames,escape = FALSE) %>%
  kable_styling(full_width = F)  %>%
      collapse_rows(columns = 1,valign = "middle")  %>%
      footnote(number = c("Model 1: fixed order by percepction/preferenc"))
sumtable01

	# CFA with secon order unmerit

	  cfa_mer_red2 <- '
	  # latent variables
		  merit =~ hwork + ambition 
		  parent=~ wealthy + pareduc
		  backg=~ race + gender
			network=~ people + polcone
		  bmerit =~ welljob + hardjob 
		  need=~ family + child
			redist =~ income_dif + unemployed + taxes 
	  	accum  =~ health + educ
			unmerit =~ parent + backg + network
	  '     
	  fit_mer_red2 <- cfa(cfa_mer_red2,data=merit_redist, 
	  	missing = "ML") 
#	  summary(fit_mer_red2, fit.measures=TRUE,standardized=TRUE)

	sum_fit<- bind_rows(fitmeasures(fit_mer_red2)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")])

sum_fit$mod <- c("Model 1")
sum_fit$est <- c("MLR")
#sum_fit <- select(sum_fit,mod,est,everything())
colnames <- c("Model","Estimator","$\\chi^2$","df","CFI","CFI.sca","CFI.rob","RMSEA","RMSEA.sca","RMSEA.rob")

sumtable01<- kable(sum_fit,digits = 3,format = "html",row.names = F,booktabs=T, caption = "Summary fit indices",col.names = colnames,escape = FALSE) %>%
  kable_styling(full_width = F)  %>%
      collapse_rows(columns = 1,valign = "middle")  %>%
      footnote(number = c("Model 1: fixed order by percepction/preferenc"))
sumtable01  
	    # Factor scores
	  fscores <- as.data.frame(lavPredict(fit_mer_red2))

	    # Correlations
	  cormat=cor(fscores[c(7,8,1,9,5,6)],
	             use="pairwise.complete.obs")
	  corrplot.mixed(cormat)
	  #save
	  dev.copy(png,"output/results/corplot_totscores.png",width=600,
	           height=600); dev.off()

	   	# Path
	 	#	  semPlot::semPaths(fit_mer_red2,layout = "tree", rotation = 2, 
		#	  	intercepts = FALSE, style = "lisrel", curvePivot = TRUE,
		#	  	sizeMan = 4, sizeLat = 6)
	 	#	  dev.copy(png,"output/results/path_mer_red2.png",width=800,
		#    	height=650);  dev.off()

	# Structural model
	  sem_mer_red1 <- '
	  # latent variables
		  merit =~ hwork + ambition 
		  parent=~ wealthy + pareduc
		  backg=~ race + gender
			network=~ people + polcone
		  bmerit =~ welljob + hardjob 
		  need=~ family + child
			redist =~ income_dif + unemployed + taxes 
	  	accum  =~ health + educ
			unmerit =~ parent + backg + network
		# regressions
			redist + accum ~ merit + unmerit + bmerit + need
	  '

	  fitsem_mer_red1 <- sem(sem_mer_red1,data=merit_redist, 
	  	missing = "ML") 
#	  summary(fitsem_mer_red1, fit.measures=TRUE,standardized=TRUE)

	  	# Path
	 		#  semPlot::semPaths(fitsem_mer_red1,layout = "tree", rotation = 2, 
			#  	intercepts = FALSE, style = "lisrel", curvePivot = TRUE,
			#  	sizeMan = 4, sizeLat = 6)
	 		#  dev.copy(png,"output/results/path_mer_red2.png",width=800,
		  #  	height=650);  dev.off()

	# saque este argumento del codigo de abajo  op=="~"

	 	# Table
	 	tabreg=subset(standardizedsolution(fitsem_mer_red1)	,op=="~", select=c(lhs,rhs,est.std,z)) 
	 	tabreg2=subset(tabreg, select=c(lhs,rhs,est.std,z))
	 	tabredist=subset(tabreg2,lhs=="redist",
	 		select=rhs:z)
	 	tabaccumt=subset(tabreg2,lhs=="accum",select=est.std:z)	
	 	tabtot=cbind(tabredist,tabaccumt)
	 	tabtot2=rbind(tabtot$est.std,tabtot$z)

# End sec 4 -----------------------------

```

```{r}
# 5.2 Descriptives ---------------------- #####
	load("input/data/proc/pref.rda")

  # Adjust data
    pref_agg=aggregate(pref, by=list(pref$v5c),
    	FUN=mean, na.rm=TRUE)
    names(pref_agg)[names(pref_agg)=="Group.1"] <- "v5c"

  # Macro descriptives

### Hay que remplazar sjp.scatter, por plot_scatter #############################################################
    ### Ademas me dice que el suindice esta fuera de los limites.

  #  plot_scatter(pref_agg$gini_net,pref_agg$accum, 
  #  	dot.labels = pref_agg$v5c,
  #  	fit.line = TRUE,
  #  	show.ci = TRUE
	#   	)

#    sjp.scatter(pref_agg$gini_net,pref_agg$redist, 
 #   	dot.labels = pref_agg$v5c,
  #  	fit.line = TRUE,
   # 	show.ci = TRUE
	  # 	)

    # Sin ZAF outlier (igual, no cambia)
  #  	pref_agg2 <- pref_agg[which(pref_agg$v5c!='ZAF'), ]
	#	    plot_scatter(pref_agg2$gini_net,pref_agg2$redist, 
	#    	dot.labels = pref_agg2$v5c,
	#    	fit.line = TRUE,
	#    	show.ci = TRUE
	#	   	)

```

```{r}
# 5.3 Models ---------------------------- ######

  #  Models SEM simple

  sem_b <- '
  	# latent variables
    redist =~ income_dif + unemployed + taxes 
    accum  =~ health + educ
		redist + accum ~ univ + quintile + sstatus + gini_market'

    fit_2c <- sem(sem_b,data=pref, ordered = c("income_dif",
      "unemployed","taxes", "health", "educ"))
    #summary(fit_2c, fit.measures=TRUE,standardized=TRUE)

#cnames <- c("Factor","Indicator","Loading (MLR)","Loading (DWLS)")
#kable(left_join(x = standardizedsolution(fit_2c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),y = #standardizedsolution(fit_2c) %>% filter(op=="=~") %>% select(lhs,rhs,est.std),c("lhs","rhs")),format = #"markdown",digits = 2,col.names = cnames, caption = "Factor loadings")

sum_fit<- bind_rows(fitmeasures(fit_2c)[c("chisq.scaled","df","cfi","cfi.scaled","cfi.robust","rmsea","rmsea.scaled","rmsea.robust")])

sum_fit$mod <- c("Model 1")
sum_fit$est <- c("MLR")
#sum_fit <- select(sum_fit,mod,est,everything())
colnames <- c("Model","Estimator","$\\chi^2$","df","CFI","CFI.sca","CFI.rob","RMSEA","RMSEA.sca","RMSEA.rob")

sumtable01<- kable(sum_fit,digits = 3,format = "html",row.names = F,booktabs=T, caption = "Summary fit indices wave 01",col.names = colnames,escape = FALSE) %>%
  kable_styling(full_width = F)  %>%
      collapse_rows(columns = 1,valign = "middle")  %>%
      footnote(number = c("Model 1: fixed order by percepction/preferenc"));
sumtable01

# Multilevel (with factor scores) ################################## No funciono nada de este apartado. 
	# check simple regresion
	#  ols_prefred=lm(redist ~ univ + quintile + sstatus + gini_market)
	#  summary(ols_prefred) # ok, similar SEM, scores OK

  # Multilevel models redist

	  # Null
	 # 	mlprefred_0 = lmer(redist ~ 1 + (1 | v5c))
	#		screenreg(mlprefred_0) # de library texreg  
	#		
#			# Graph
#			sjp.lmer(mlprefred_0, sort.est="(Intercept)")
#	  
			# ICC
#			varcomp=as.data.frame(VarCorr(mlprefred_0))
#			tau00=varcomp[1,4]
#			sigma2=varcomp[2,4]
#			tau00/(tau00+sigma2)

		# Individual level vars
#			mlprefred_1 = lmer(redist_n ~ 1 + univ + 
#					as.factor(quintile_miss) +	sstatus + sex + age + 
#					(1 | v5))
#			screenreg(mlprefred_1)
#			
#		# Macro level vars
#			mlprefred_2 = lmer(redist ~ 1 + univ + 
#					as.factor(quintile_miss)+ sstatus + gini_market + 
#					univ_g + (1 | v5))
#			screenreg(mlprefred_2)
#			
#	# Multilevel models accumulation
#			
	  # Null
#	  	mlaccum_0 = lmer(accum ~ 1 + (1 | v5c))
#			screenreg(mlaccum_0) # de library texreg  
#	  
			# Graph
#			sjp.lmer(mlaccum_0, sort.est="(Intercept)")
#			
			# ICC
#			varcomp=as.data.frame(VarCorr(mlaccum_0))
#			tau00=varcomp[1,4]
#			sigma2=varcomp[2,4]
#			tau00/(tau00+sigma2)
#		
#		# Individual level vars
#			mlaccum_1 = lmer(accum ~ 1 + univ + as.factor(quintile_miss) +	
#					sstatus + sex + age + univ*sex + age*age + (1 | v5))
#			screenreg(mlaccum_1)
#			
		# Macro level vars
#			mlaccum_2 = lmer(accum ~ 1 + univ_g + as.factor(quintile_miss) +
#					(sstatus -f_sstatus) + gini_net + (1 | v5))
#			screenreg(mlaccum_2)

		# Random slope & cross-level
#			mlaccum_3 = lmer(accum ~ 1 + univ_g + f_sstatus + gini_net + 
#					gini_net* f_sstatus +	(1 + f_sstatus | v5))

#			screenreg(mlaccum_3)

#			# Graphs
#			sjp.lmer(mlaccum_3, sort.est = "(Intercept)")
#			sjp.lmer(mlaccum_3, type = "rs.ri")
#			
#			mlaccum_4 = lmer(accum ~ 1 + univ + f_sstatus + gini_net + gini_net*f_sstatus + hhincome +	(1 + hhincome | v5))
#			screenreg(mlaccum_4)

# End sec. 5 ----------------------------

```

# 6. MODELS C: MERITOCRACY ================ =

```{r}

# 6.2 Descriptives ---------------------- #####

    load("input/data/proc/meritocracy.rda")

  # Adjust data
    merit_agg=aggregate(meritocracy, by=list(meritocracy$v5c),
    	FUN=mean, na.rm=TRUE)
    names(merit_agg)[names(merit_agg)=="Group.1"]="v5c"
 #   names(merit_agg)
  #  dim(merit_agg)

  # Macro descriptives

    ################################agregar Polichoric.
    # Macro level corr

  #    corMat <- cor(merit_agg[c("merit","unmerit","bmerit","need","gini_net","gini_market","j_fi_index")], use="pairwise.complete.obs") # estimar matriz pearson
  #    options(digits=2)
  #    corMat # muestra matriz

  #    M=cor(merit_agg) # matriz simple 
  #    corrplot.mixed(corMat)

      ##### la mayoria de los str.scatter precentan problemas en los limites "subindice fuera de limites.

  #  sjp.scatter(merit_agg$gini_net,merit_agg$bmerit, 
  #  	dot.labels = merit_agg$v5c,
  #  	fit.line = TRUE,
  #  	show.ci = TRUE
	#   	)
  #    # Save
  #    dev.copy(png,"output/results/gini_bmerit.png"); dev.off()
  #

  #  sjp.scatter(merit_agg$gini_net,merit_agg$unmerit, 
   # 	dot.labels = merit_agg$v5c,
    #	fit.line = TRUE,
  #  	show.ci = TRUE
	#   	)    
# 
 #  sjp.scatter(merit_agg$gini_net,merit_agg$need, 
  #  	dot.labels = merit_agg$v5c,
   # 	fit.line = TRUE,
    #	show.ci = TRUE
	   #	)

         # Save
      #dev.copy(png,"output/results/fi_need.png"); dev.off()

```

![Imagen provisoria](../output/results/fi_need.png)

```{r}
   #sjp.scatter(merit_agg$j_fi_index,merit_agg$need, 
    #	dot.labels = merit_agg$v5c,
  #  	fit.line = TRUE,
   # 	show.ci = TRUE
	  # 	)     
      # Save
     # dev.copy(png,"output/results/gini_fi.png"); dev.off()
```

![Imagen provisoria](../output/results/gini_fi.png)

```{r}
  # Sin outliers

  	merit_agg2 <- merit_agg[which(merit_agg$v5c!="ZAF" & merit_agg$v5c!="CHN" &  merit_agg$v5c!="CHL" ), ]

    #sjp.scatter(merit_agg2$gini_net,merit_agg2$unmerit, 
    #	dot.labels = merit_agg2$v5c,
    #	fit.line = TRUE,
    #	show.ci = TRUE
	 #  	)

```

```{r results='asis'}
# 6.3 Models -----------------------------#####

# Multilevel (with factor scores)

  load("input/data/proc/meritocracy.rda")

    # Adjust some variables
    	meritocracy$Quintile=as.factor(meritocracy$quintile_miss)
     	levels(meritocracy$Quintile)[6] <- "miss"

	# check simple regresion
#	  ols_merit=lm(merit ~ univ + Quintile + sstatus + gini_net, data = meritocracy)
#	  summary(ols_merit) # ok, similar SEM, scores OK

  # Multilevel models merit

	  # Null
	  	mlmerit_0 = lmer(merit ~ 1 + (1 | v5c), data = meritocracy)
			screenreg(mlmerit_0) # de library texreg

			# Graph
		#	sjp.lmer(mlmerit_0, sort.est="(Intercept)")

			# ICC
			varcomp=as.data.frame(VarCorr(mlmerit_0))
			tau00=varcomp[1,4]
			sigma2=varcomp[2,4]
			tau00/(tau00+sigma2)

  		# Individual level vars
  			mlmerit_1 = lmer(merit ~ 1 + univ +	quintile_miss +	
  					sstatus + sex + age + (1 | v5), data = meritocracy)
  			screenreg(mlmerit_1)

		# Macro level vars
			mlmerit_2 = lmer(bmerit ~ 1 + gini_net + (1 | v5), data = meritocracy)
			screenreg(mlmerit_2)

# Models meritocracy and preferences

  load("input/data/proc/meritocracy.rda")
  load("input/data/proc/redistrib_sco.rda")

  	mer_pref=merge(meritocracy, redistrib_sco, by=c("id"))
 #   names(mer_pref)

  mer_pref_agg=aggregate(mer_pref, by=list(mer_pref$v5c),
    	FUN=mean, na.rm=TRUE)
      names(mer_pref_agg)[names(mer_pref_agg)=="Group.1"]="v5c"
	 #  names(mer_pref_agg)
  #  dim(mer_pref_agg)

    # Graphs    # mismo problemas de subindices.

  #    plot_scatter(mer_pref_agg$merit,mer_pref_agg$bmerit, 
   # 	dot.labels = mer_pref_agg$v5c,
    #	fit.line = TRUE,
  #  	show.ci = TRUE
	 #  	)
  #    
  #    plot_scatter(mer_pref_agg$merit,mer_pref_agg$redist, 
  #  	dot.labels = mer_pref_agg$v5c,
   # 	fit.line = TRUE,
    #	show.ci = TRUE
	 #  	)
      # Save
  #    dev.copy(png,"loc_merit_pref_int/results/merit_redis.png"); dev.off()

    #  sjp.scatter(mer_pref_agg$unmerit,mer_pref_agg$redist, 
  #  	dot.labels = mer_pref_agg$v5c,
   # 	fit.line = TRUE,
    #	show.ci = TRUE
	   	#)      
       # Save
     # dev.copy(png,"loc_merit_pref_int/results/unmerit_redis.png"); dev.off()

  #    sjp.scatter(mer_pref_agg$redist,mer_pref_agg$accum, 
  #  	dot.labels = mer_pref_agg$v5c,
  #  	fit.line = TRUE,
  #  	show.ci = TRUE
	#   	)

```

![IMAGEN PROVISORIA](../output/results/merit_redis.png)