### 2020-07-08 production/prod_prep-data.rmd

---
title: "Data preparation"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
css: "../input/bib/style.css"
output:
  html_document:
    number_sections: false
    toc: true
    toc_float:
        collapsed: true
    toc_depth: 2
    code_folding: "hide"
editor_options:
  chunk_output_type: console
---

```{r eval=FALSE, include=FALSE}
rmarkdown::render(input = "production/prod_prep-data-A.Rmd",output_format = "html_document",output_dir = "production/");browseURL(url = "production/prod_prep-data-A.html")
```

```{r include=FALSE}
knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = FALSE,results = "hold")
knitr::opts_knit$set(root.dir = "../")
options(scipen=999)
rm(list=ls())
options(knitr.kable.NA = '')
```

# Document content: Data preparation

In this site the production and adjustment of the databases that are used for the multilevel analysis are exposed. The work in general focuses on selecting and recoding variables obtained from international databases such as the International Social Survey (ISSP) and the Standardized World Income Income Inequality Database (SWIID). In addition, factor analyzes are performed to subsequently impute predicted values that represent the estimated latent variables.

In diagram 1, a summary of the data processing is presented, which starts from international databases (input) to processed databases. As shown in the graph, the document is divided into 4 sections , corresponding to the subtitles of the codes.

In the first section, in addition to selecting and recoding ISSP variables, factor analyzes are performed to evaluate and create the latent variables of preferences about redistribution and inequality. In section 2, international SWIID and GOQ data are collected and added together with ISSP data, which includes predicted values for latent variables of meritocratic perceptions and preferences. In the next two sections, adjustments are made for multilevel analyzes for the pref.rda and meritocracy.rda databases.

### load packages
```{r}
library(sjPlot)
library(sjmisc) # several functions,ej read_stata
library(descr)
library(memisc)  # codebook
library(psych) # cor2latex
library(corrplot) # graph correlations
library(lavaan)
library(stargazer)
library(semTools)
library(plyr) # count
library(dplyr)
#library(car) # recode
library(ggplot2)
#library(Hmisc)
library(tidyverse)
library(data.table)
library(stringr)
library(countrycode)
library(lme4)
library(texreg)
library(mice)
library(ggrepel) # for scatter labels in sjp.scatter
#library(semPlot)
library(haven)
library(sjlabelled)
library(kableExtra)
library(summarytools)
library(tidyselect)
library(magrittr)
library(semPlot)
#library(Hmisc)
library(knitr)

```

# Section-1: Preparation individual variables ISSP

## General preparation
```{r results='hide'}
load(file = "input/data/original/issp.rda")
#load(url('https://raw.githubusercontent.com/franciscomeneses/bases/master/issp.rda'))

    # Check data  
    #names(issp2009)

### change variable names to lowercase
names(issp2009) <- tolower(names(issp2009))

### Check & generate IDs (v4: respondent, v5: country)
### Generate new sequential id individuals by country to merge
      issp2009$id <- sequence(tabulate(issp2009
                                     $v5))
    issp2009$idtem=issp2009$v5*10000  
    issp2009$id=issp2009$idtem + issp2009$id
```

##  Preparation-data of the redistributive preference scales: redistrib.rda

| Code | New Code   | Variable                                                                                                        |
|------|------------|-----------------------------------------------------------------------------------------------------------------|
| v33  | income_dif | Gov. resp to reduce differences in income                                                                       |
| v34  | unemployed | Gov. resp to provide decent s andard of living for unemployed                                                   |
| v36  | taxes      | People with high income should pay a larger / same smaller sha3re of income in taxes than those with low income |
| v38  | health     | ok if people with high income can buy better health                                                             |
| v39  | educ       | ok if people with high income can buy better education                                                          |                     Table:`Items for latent variables`

###  Rename & define object

```{r results='asis'}
# select
    redistrib=issp2009[,c("v5","v33","v34","v36","v38","v39","id")]

# new names(issp2009)
    redistrib$income_dif= redistrib$v33
    redistrib$unemployed= redistrib$v34
    redistrib$taxes     = redistrib$v36
    redistrib$health    = redistrib$v38
    redistrib$educ      = redistrib$v39

#Delete old variables

    redistrib=redistrib[-c(2:6)]

    stargazer(redistrib[-c(1,2)], type="html")  
```

### Recode NAs
```{r results='asis'}
    redistrib[3:7][redistrib[3:7] > 5] <- NA
   stargazer(redistrib[-c(1,2)], type="html")

```

```{r}
dim(redistrib)
```

### adjust labels NAs
```{r}
     get_labels(redistrib$income_dif)
    redistrib[,3:7] <-sjlabelled::set_na(redistrib[,c("income_dif","unemployed", "taxes", "health", "educ")], na = c(8,9))

```

### Reverse code for SEM models
```{r}
redistrib[3:7]=6-redistrib[3:7]

```

### 2.4 Factor scores

```{r cache=TRUE, results='hide'}
cfa_red1 <- '
	# latent variables
	redist =~ income_dif + unemployed + taxes 
	accum  =~ health + educ
	'

	  # generate 5 multiple complete datasets 
		out <- mice(redistrib[3:7], m=5)

```

```{r include=FALSE}

		D1 <- complete(out, 1) 
		D2 <- complete(out, 2) 
		D3 <- complete(out, 3) 
		D4 <- complete(out, 4) 
		D5 <- complete(out, 5)

		# fit model for each complete dataset 
		fit1 <- cfa(cfa_red1, data=D1) 
		fit2 <- cfa(cfa_red1, data=D2) 
		fit3 <- cfa(cfa_red1, data=D3) 
		fit4 <- cfa(cfa_red1, data=D4) 
		fit5 <- cfa(cfa_red1, data=D5)

	# predict scores for all models 
		p1 <- predict(fit1) 
		p2 <- predict(fit2) 
		p3 <- predict(fit3) 
		p4 <- predict(fit4) 
	p5 <- predict(fit5)

# compute 'average' across 5 sets of scores: 
		scores <- (p1 + p2 + p3 + p4 + p5)/5

 #Generate factor scores      
   scores_dat=as.data.frame(scores)

# Adaptar base para agregar factor scores (sacar todos NA)  
   redistrib_m=redistrib[rowSums(is.na(redistrib[3:7]))!=5,]

# Merge with factor scores
	  redistrib_sco=cbind(redistrib,scores)

	  # Check
	   stargazer(redistrib_sco[,8:9], type ="text")

	# Save  factor scores
	  save(redistrib_sco, file="input/data/proc/redistrib_sco.rda")
```

redistrib_sco [56021,9]

```{r}
print(dfSummary(redistrib_sco, valid.col = FALSE, graph.magnif = 1),
      max.tbl.height = 300, method = "render")
```

## Individual level variables: Sociodemograficas.

```{r results='asis'}
### Select

indepvar= issp2009 %>% dplyr::select(contains("_inc"), "sex","age","degree", "hompop", "v44", "v45", "movstatus"="v46", "v5","id")

### Recode sex

sjPlot::plot_frq(indepvar$sex, title ="Gender before")
indepvar$sex=ifelse(indepvar$sex==1,0,ifelse(indepvar$sex==2,1,NA))  
sjPlot::plot_frq(indepvar$sex, title ="gender after")

### Recode Age

  # Age
sjPlot::plot_frq(indepvar$age, title = "Age before",show.mean = T , show.values = F,show.n = F)

#which(indepvar$age > 85) # / 461 casos !
#which(indepvar$age < 15) # / 0 casos !
indepvar$age[indepvar$age > 85] = NA

# Categorical age
indepvar$age_c=indepvar$age
indepvar$age_c[indepvar$age_c <= 25] = 1
indepvar$age_c[indepvar$age_c > 25 & indepvar$age_c <=35] = 2
indepvar$age_c[indepvar$age_c > 35 & indepvar$age_c <=45] = 3
indepvar$age_c[indepvar$age_c > 45 & indepvar$age_c <=55] = 4
indepvar$age_c[indepvar$age_c > 55 & indepvar$age_c <=65] = 5
indepvar$age_c[indepvar$age_c > 65] = 6

sjPlot::plot_frq(indepvar$age_c, title = "Age after", show.na = T)

### Recode Education

sjPlot::plot_frq(indepvar$degree, coord.flip = TRUE, title = "Degree before")

  # Recode. Values over 5 are removed

  indepvar$degree[indepvar$degree > 5]=NA

#### Dummy university
# Dummy university
  indepvar$univ=indepvar$degree
  indepvar$univ=ifelse(indepvar$univ<=4,0,
  ifelse(indepvar$univ==5,1,NA))

  sjPlot::plot_frq(indepvar$univ, coord.flip = TRUE, title = "Dummy university")

### Recode Income

  stargazer(indepvar[1:41], type="html")

  # Income (variables finish in _inc)
# Continuo

   # Recode missings
      # Para todos los paises, recode missings.  
      indepvar[1:41][indepvar[1:41] >= 9999990] <- NA
      indepvar[1:41][indepvar[1:41] == 999990] =NA
      indepvar[1:41][indepvar[1:41] == 999997] =NA
  indepvar[1:41][indepvar[1:41] == 999998] =NA
  indepvar[1:41][indepvar[1:41] == 999999] =NA

    # Identificar ceros (para poder diferenciar de los NAs luego)
    indepvar[1:41][indepvar[1:41] == 0] <- 111111

  # Agregar ahora todos en un solo vector
indepvar$income=rowSums(indepvar[1:41],na.rm = TRUE)

indepvar$income[indepvar$income ==0]=NA
indepvar$income[indepvar$income ==1111]=0

stargazer(indepvar[1:41], type="html")

#summary(indepvar$income)

### Recode Household income

# Household income

# Preparar numero de personas en household
indepvar$hompop[indepvar$hompop==99]=NA

# Ahora dividir el vector income por n° de personas en hogar
indepvar$hhincome=indepvar$income/indepvar$hompop
summary(indepvar$hhincome)

# Y ajustar NAs de la variable final ingreso continuo
indepvar$hhincome[indepvar$hhincome==0]= NA
indepvar$hhincome[indepvar$income==111111]= 0
indepvar$hhincome[indepvar$hompop==0]= NA

#Wsummary(indepvar$hhincome)

# check
income=indepvar[c("v5", "ar_inc", "hompop","income",
"hhincome")]
# View(income)

# Generar quintiles de ingreso
indepvar %>% group_by(v5) %>% mutate(quintile =
ntile(hhincome,5)) ->indepvar

# Check
#prop.table(table(indepvar$v5,indepvar$quintile),1)*100

# Generar quintiles con NA como categorÃ?a
indepvar$quintile_miss=indepvar$quintile
indepvar$quintile_miss[is.na(indepvar$quintile_miss)] = 6

sjPlot::plot_frq(indepvar$quintile_miss, coord.flip = TRUE, title = "Quintiles de ingreso")
#stargazer(indepvar$quintile_miss, type = "html")

### Recode Subjetive status

  # Subjective status

# Individual
#freq(indepvar$v44)
indepvar$sstatus=indepvar$v44
indepvar$sstatus[indepvar$ssttus>10]=NA
  #freq(indepvar$sstatus)

    # Family
#freq(indepvar$v45)
indepvar$f_sstatus=indepvar$v45
    indepvar$f_sstatus[indepvar$f_sstatus>10]=NA
  #freq(indepvar$f_sstatus)

    knitr::kable(table(indepvar$movstatus), "html")

    indepvar$movstatus[indepvar$movstatus>=8]= NA

    knitr::kable(table(indepvar$movstatus), "html")

```

```{r}
#Codigos por incorporar: se deben hacer ajustes de moneda y partidos por pais.

# Home value (ver Nota 3)
#summary(indepvar$v67)

  # Party
    #freq(issp2009$party_lr)
#prop.table(table(indepvar$v5,indepvar$party_lr),1)*100
# paises sin datos: CL, CN, CY, HU, IL, TW
#freq(indepvar$cl_prty) # estan
#freq(indepvar$cn_prty) # no
#freq(indepvar$cy_prty) # si
#freq(indepvar$hu_prty) # si
#freq(indepvar$il_prty) # si
#freq(indepvar$tw_prty) # si

# para usarlas habría que clasificar partidos de esos países

```

```{r}
#Eliminar income por paises. 
indepvar=indepvar[-c(1:41)]
dim(indepvar)
View(indepvar)
```

```{r}
#indepvar_graph<-as.data.frame(indepvar[,2:16])

print(dfSummary(indepvar[-c(8)], valid.col = FALSE, graph.magnif = 1),
      max.tbl.height = 300, method = "render")
```

## Measurement Merit & Getting ahead

| Code  	| New Code  	| Variable              	|
|-------	|-----------	|-----------------------	|
| v6    	| wealthy   	| wealthy family        	|
| v7    	| pareduc   	| educated parents      	|
| v9    	| ambition  	| ambition              	|
| v10   	| hwork     	| hard work             	|
| v11   	| people    	| right people          	|
| v12   	| polcone   	| political connections 	|
| v14   	| race      	| race                  	|
| v16   	| gender    	| gender                	|
| v5    	| -         	| Country               	|
| id    	| -         	| inividual id          	|

```{r results='hide'}
    # Rename
    getahead=issp2009[,c("v6","v7","v9", "v10","v11","v12",
             "v14","v16", # get ahead
    	  		 "v5","id")]

    getahead$wealthy=getahead$v6
    getahead$pareduc=getahead$v7
    getahead$ambition= getahead$v9
    getahead$hwork=getahead$v10
    getahead$people=getahead$v11
    getahead$polcone=getahead$v12
    getahead$race=getahead$v14
    getahead$gender=getahead$v16

    getahead=getahead[-c(1:8)]

  # recode values 8 9 to NA
    #freq(getahead$wealthy)
    getahead[3:10][getahead[3:10] > 5] <- NA
    #freq(getahead$gender) # check ok
    stargazer(getahead, type = "html")

  # adjust labels
    get_labels(getahead$wealthy)
    getahead[ , 3:10][getahead[ , 3:10] >= 8] <- NA
    get_labels(getahead[3:10]) # check

  # Reverse code for SEM models
    getahead[3:10]=6-getahead[3:10]

  # Generate variables with 3 values for getahead mess model categoric
    #names(getahead)
    getahead_temp=subset(getahead[c(2,3:10)])
    #names(getahead_temp)
    getahead_temp[2:9][getahead_temp[2:9] <  4] <- 3
   # freq(getahead_temp$hwork)
    #freq(getahead$hwork)

    colnames(getahead_temp)=paste("c3", colnames(getahead_temp),
    	sep="")
    names(getahead_temp)[names(getahead_temp)=="c3id"] <- "id"
    getahead_temp2=merge(getahead,getahead_temp, by=c("id"))
  	names(getahead_temp2) # ok
  	getahead=getahead_temp2

```

```{r}
plot_likert((getahead[,3:8]), 
                        geom.colors   = c("#9ecae1", "#6baed6","#4292c6", 
                                            "#2171b5"), # colorbrewer2.org 
                        sort.frq      = "neg.asc", # sort descending)
                        title         = "getahead",
                        intercept.line.color = "white", # vertical middle
                        expand.grid   = F, # no inner margins in plot
                        show.n        = FALSE, # hide N's in axis labels
                        grid.range    = 1.8,
                        geom.size = 0.6) # expand axis 130% )
```

```{r}
	# Save object
  #  save(getahead, file="../objects/getahead.rda")

```

### Detearn: Data preparation, detearn.rda ----- #####

```{r}
		#names(issp2009)
	  detearn=issp2009[,c("v47", "v48", "v49", "v50", "v51",
	   	"v52", "v5","id")]
names(detearn)
    detearn$respons 	= detearn$v47
    detearn$yeduc  	  = detearn$v48
    detearn$family    = detearn$v49
    detearn$child     = detearn$v50
    detearn$welljob   = detearn$v51
    detearn$hardjob   = detearn$v52
names(detearn)
    detearn=detearn[-c(1:6)]
names(detearn)
  # recode values 8 9 to NA
    summary(detearn)
  # freq(detearn$respons)
    detearn[3:8][detearn[3:8] > 5] <- NA

      # freq(detearn$respons) # check ok

    knitr::kable(summary(detearn), "html")

  # Out Portugal (no answer in well job item)
    detearn[which(detearn$v5==620),]=NA
  # adjust labels ###PROB PORTUCAL!!!
    get_labels(detearn)
detearn[ , 3:8][detearn[ , 3:8] >= 8] <- NA
    detearn$welljob<-sjlabelled::set_na(detearn$welljob, na =c(0))
        knitr::kable(summary(detearn), "html")
   # freq(detearn$welljob) ## prob portugal
    detearn$welljob[detearn$welljob ==0] = NA
    get_labels(detearn$welljob)
    labels(detearn$welljob)=labels(detearn$respons)### PROB NOT SOLVED

# Reverse code for SEM models
    detearn$respons <- as.numeric(detearn$respons)
    detearn$yeduc <- as.numeric(detearn$yeduc)
    detearn$family <- as.numeric(detearn$family)
    detearn$child <- as.numeric(detearn$child)
    detearn$welljob <- as.numeric(detearn$welljob)
    detearn$hardjob <- as.numeric(detearn$hardjob)

```

```{r}
    sjPlot::plot_likert(detearn[,3:8],
                        geom.colors   = c("#9ecae1", "#6baed6","#4292c6", "#2171b5"), # colorbrewer2.org 
                        cat.neutral = 3,
                        sort.frq      = "neg.asc", # sort descending)
                        title         = "Before detearn",
                        intercept.line.color = "white", # vertical middle
                        expand.grid   = F, # no inner margins in plot
                        show.n        = FALSE, # hide N's in axis labels
                        grid.range    = 1.8,
                        geom.size = 0.6) # expand axis 130% ))

detearn$respons = as.numeric(6 - detearn$respons)
detearn$yeduc = as.numeric(6 - detearn$yeduc)
detearn$family = as.numeric(6 - detearn$family)
detearn$child = as.numeric(6 - detearn$child)
detearn$welljob = as.numeric(6 - detearn$welljob)
detearn$hardjob = as.numeric(6 - detearn$hardjob)

  sjPlot::plot_likert(detearn[,3:8],
                        geom.colors   = c("#9ecae1", "#6baed6","#4292c6", "#2171b5"), # colorbrewer2.org 
                      cat.neutral = 3,
                        sort.frq      = "neg.asc", # sort descending)
                        title         = "After detearn",
                        intercept.line.color = "white", # vertical middle
                        expand.grid   = F, # no inner margins in plot
                        show.n        = FALSE, # hide N's in axis labels
                        grid.range    = 1.8,
                        geom.size = 0.6) # expand axis 130% ))

freq(detearn$respons)
```

## Measuring Beliefs & Perceptions: merit_pref_perc

```{r}
# 3. MEASURING BELIEFS & PERCEPTIONS ------ -

# 3.1 Define object --------------------- -------------------------------

#			load("./objects/getahead.rda")
#			load("./objects/detearn.rda")
		#	names(getahead)
		#	names(detearn)

			merit_per_pref=merge(getahead,detearn, by=c("id","v5"))
		#	names(merit_per_pref)

  # Estimación del modelo c/ unmerit 2do orden
	  cfa_perpref2 <- '
	  # latent variables
	  merit =~ hwork + ambition
	  parent=~ wealthy + pareduc
	  backg=~ race + gender
		network=~ people + polcone
		unmerit=~ parent + backg + network # 2nd order
	  bmerit =~ welljob + hardjob
	  need=~ family + child
	  '
	  fit_perpref2 <- cfa(cfa_perpref2,data=merit_per_pref,
	  	missing = "ML")
#	  summary(fit_perpref2, fit.measures=TRUE,standardized=TRUE)

  # Path model
		  semPaths(fit_perpref2,layout = "tree", rotation = 2,
		  	intercepts = FALSE, style = "lisrel", curvePivot = TRUE,
		  	sizeMan = 3.5, sizeLat = 6)
```

### Factor scores fit_perpref2: merit_per_pref_sco

```{r cache=TRUE, results='hide'}

			out <- mice(merit_per_pref[,c(3:10,19:24)], m=5)
```

```{r results='hide'}
		# Multiple imputation (para recuperar missings en scores) - REVIEW THIS, command lavPredict from 2017 should predict with missing values therefore saving from multiple imputation

		  # generate 5 multiple complete datasets

			D1 <- complete(out, 1)
			D2 <- complete(out, 2)
			D3 <- complete(out, 3)
			D4 <- complete(out, 4)
			D5 <- complete(out, 5)

			# fit model for each complete dataset
			fit1 <- cfa(cfa_perpref2, data=D1)
			fit2 <- cfa(cfa_perpref2, data=D2)
			fit3 <- cfa(cfa_perpref2, data=D3)
			fit4 <- cfa(cfa_perpref2, data=D4)
			fit5 <- cfa(cfa_perpref2, data=D5)

		# predict scores for all models
			p1 <- predict(fit1)
			p2 <- predict(fit2)
			p3 <- predict(fit3)
			p4 <- predict(fit4)
			p5 <- predict(fit5)

		# compute 'average' across 5 sets of scores:
			scores <- (p1 + p2 + p3 + p4 + p5)/5

	  # Generate factor scores
	    scores_dat=as.data.frame(scores)
	    dim(scores_dat)
	    str(scores_dat)
	    summary(scores)

		# Generate object with factor scores
	    dim(merit_per_pref)
	    dim(scores_dat)
		  merit_per_pref_sco=cbind(merit_per_pref,scores)

	# Save object with factor scores
	#	names(merit_per_pref_sco)
		merit_per_pref_sco=merit_per_pref_sco[,c(1:2,25:31)]
#	  save(merit_per_pref_sco,
#	  	file="./objects/merit_per_pref_sco.rda")

```

Bases de datos creadas.

[Explicar mejor]

Indepvar.rda: variables sociodemograficas
redistrib_sco: variables latentes de preferencias redistributivas.
merit_perc_pref_sco: variables latentes de creencias meritocraticas.

## Merge varaibles indivuales.

```{r}

indiv<- merge (indepvar, redistrib_sco, by = c("id"))

indiv <- merge (indiv,  merit_per_pref_sco, by = c("id"))
```

```{r}
print(dfSummary(indiv, valid.col = FALSE, graph.magnif = 1),
      max.tbl.height = 300, method = "render")
```

```{r}
save(indiv, file="input/data/proc/indiv.rda")
```

# Section 2: agregar datos

```{r}
agg_lm=aggregate(indiv, by=list(indiv$v5),FUN=mean, na.rm=TRUE)

colnames(agg_lm) <- paste("j", colnames(agg_lm), sep = "_")
table(agg_lm$j_v5)

agg_lm$v5=agg_lm$j_v5

mer_pref=merge(indiv,agg_lm, by="v5")

 dim(mer_pref)

```

```{r}
    mer_pref$age_g=mer_pref$age-mer_pref$j_age
    mer_pref$sex_g =mer_pref$sex-mer_pref$j_sex
    mer_pref$univ_g =mer_pref$univ-mer_pref$j_univ
  mer_pref$sstatus_g= mer_pref$sstatus-mer_pref$j_sstatus

    summary(mer_pref$univ_g)
```

# Seccion 3:  add macro var

## Section 2.1: SWIID
```{r}
# Inequality swiid (4)        
#load(url("https://raw.githubusercontent.com/franciscomeneses/bases/master/swiid8_3.rda"))
load("input/data/original/swiid8_3.rda")

# convertir desde formato original list a dataframe

swiid_summary <- swiid %>%  bind_rows() %>%
group_by(country, year) %>%  
summarize_all(funs(mean, sd)) %>%
ungroup() %>% dplyr::rename(.dots=setNames(names(.),
str_replace(names(.), "_mean", ""))) %>%  
dplyr::rename_(.dots=setNames(names(.),str_replace(names(.), "_sd", "_se")))

save(swiid_summary, file = "input/data/proc/swiid_summary.rda")

```

###  Filter only 2009
```{r}
summary(swiid_summary$.dots...year)

swiid_summary09=subset(swiid_summary, .dots...year==2009)
```

###  Merge with indepvar object

```{r}
# Generate equivalente country code

head(swiid_summary09$.dots...country)
as.numeric(swiid_summary09$country)

# Continuo
swiid_summary09$v5=countrycode(swiid_summary09$.dots...country,'country.name','iso3n')
str(swiid_summary09$v5)

# Categórico
swiid_summary09$v5c=countrycode(swiid_summary09$.dots...country,'country.name','iso3c')
str(swiid_summary09$v5c)

# Merge
#indepvar <- merge(swiid_summary09,indepvar,by=c("v5"))

#summary(indepvar$.dots..._1_gini_disp)

```

```{r}
mer_pref<- merge(swiid_summary09,mer_pref, by=c("v5"))
```

## Section 2.2: Qog

```{r echo=FALSE, results='hide'}
#Para descargar

#devtools::install_github("ropengov/rqog")
#library("rqog")

#qog2009_dta<-read_qog(
#  which_data = "standard",
#  data_type = "time-series",
#  year = 2009,
#  data_dir = NULL,
#  file_format = "csv",
#  download_only = FALSE,
#  cache = TRUE,
#  update_cache = FALSE
#)

```

```{r results='hide'}
load( file = "input/data/original/qog.rda")
#load(url('https://raw.githubusercontent.com/franciscomeneses/bases/master/qog.rda'))

		# Extract 2009 data
			qog2009=subset(qog, year==2009)

		# Merge with ISSP countries for cleaning

			load("input/data/proc/indepvar_contries.rda")
			issp_countries=	aggregate(indepvar_contries[c("v5")], 
					list(indepvar_contries$country,indepvar_contries$v5c),mean)

			qog2009[c("ccode","cname")] # check ISO codes
			names(qog2009)[names(qog2009) == 'ccode'] <- 'v5'
			qog2009_c=merge(issp_countries, qog2009, by=c("v5")) 
			dim(qog2009_c)
			colSums(is.na(qog2009_c))

			# Eliminar columnas con máximo 5 NAs
			qog2009_c_na= qog2009_c[ , colSums(is.na(qog2009_c)) < 6]

			qog2009_c_na= qog2009_c_na %>% dplyr::select("v5", "fi_index") %>% as.data.frame()
			#names(qog2009_c_na)

```

# Merge
```{r}

mer_pref <- merge(mer_pref, qog2009_c_na, by=c("v5"))
```

End sec 2 objeto: mer_pref.rda -------

```{r}
mer_pref$gini_neto<- mer_pref$.dots...gini_disp
mer_pref$gini_mkt <- mer_pref$.dots...gini_mkt
```

```{r}
save(mer_pref, file="input/data/proc/mer_pref.rda")
```

```{r}
mer_pref_polycor <- mer_pref %>% dplyr::select( 
"j_redist", "j_accum" , "j_merit","j_unmerit","j_bmerit","j_need",".dots...gini_disp",".dots...gini_mkt","fi_index")
save(mer_pref_polycor, file ="input/data/proc/mer_pref_polycor.rda")
```

mer_pref_polycor  cambiar nombre